---
title: <font size="7"><b>Spotting and measuring schroeder periods</b></font>
subtitle: <font size="4"><b>Fine scale acoustic perception in zebra finches</b><br></font>
author: <font size="4"><a href="http://marceloaray7asalas.weebly.com/">Marcelo Araya-Salas</a></font>
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: show
    css: extra.css
    df_print: tibble
    highlight: pygments  
    toc: yes
    toc_depth: 3
    toc_float:
      collapsed: yes
      smooth_scroll: yes
fontsize: 12pt 
editor_options: 
  chunk_output_type: console
---

<!-- this code add line numbers to code blocks -->
<!-- only works when code folding is not used in yaml (code_folding: show) -->

<style>
body
  { counter-reset: source-line 0; }
pre.numberSource code
  { counter-reset: none; }
</style>

```{r load packages and setup style, echo = FALSE, message = FALSE, warning=FALSE}

# github packages must include user name ("user/package")
# knitr is require for creating html/pdf/word reports
# kableExtra is used to print pretty formatted tables 
# formatR is used for soft-wrapping code
# klippy is used for adding a copy button to each code block
pkgs <- c("kableExtra", "knitr", "formatR", "rprojroot", "xaringanExtra", "warbleR", "viridis")

# install/ load packages
sketchy::load_packages(pkgs, quite = TRUE)

# set working directory as project directory or one directory above,
rootdir <- try(rprojroot::find_rstudio_root_file(), silent = TRUE)
if (is(rootdir, "try-error")) rootdir <-  ".."
opts_knit$set(root.dir = rootdir)

# options to customize chunk outputs
knitr::opts_chunk$set(
  class.source = "numberLines lineAnchors", # for code line numbers
  tidy.opts = list(width.cutoff = 65), 
  tidy = TRUE,
  message = FALSE
 )

# this is a customized printing style data frames 
# screws up tibble function
tibble <- function(x, ...) { 
  x <- kbl(x, digits=4, align= 'c', row.names = FALSE) 
   x <- kable_styling(x, position ="center", full_width = FALSE,  bootstrap_options = c("striped", "hover", "condensed", "responsive")) 
   asis_output(x)
}

registerS3method("knit_print", "data.frame", tibble)

# to add copy button to code blocks
htmltools::tagList(
  xaringanExtra::use_clipboard(
    button_text = "<i class=\"fa fa-clipboard\"></i>",
    success_text = "<i class=\"fa fa-check\" style=\"color: #90BE6D\"></i>",
    error_text = "<i class=\"fa fa-times-circle\" style=\"color: #F94144\"></i>"
  ),
  rmarkdown::html_dependency_font_awesome()
)

```

&nbsp; 

<!-- skyblue box -->

<div class="alert alert-info">

# Purpose

- Detect start and end of periods in a shroeder using amplitude cross-correlation
- Figure out a way to measure fine scale structural variation and periodicity in schroeders

</div>

&nbsp;

<!-- light brown box -->
<div class="alert alert-warning">

# Report overview

  - [Exploring data](#exploring-data)
  - [Amplitude cross-correlation](#amplitude-cross-correlation)
  - [Create a function to spot periods](#create-a-function-to-spot-periods)
  - [Test function](#test-function)
  - [Takeaways](#takeaways)

</div>

&nbsp;
&nbsp; 
```{r}

wave_col <- viridis(10)[7]

```



# Exploring data

sound file `SchroederCallsTest4_3Nov2022.wav`:
```{r}

schroeders <- read_wave("./data/raw/SchroederCallsTest4_3Nov2022.wav")

spectro(schroeders, scale = FALSE, osc = TRUE, fastdisp = TRUE, grid = FALSE, palette = viridis, collevels = seq(-100, 0 , 5), wl = 512, colwave= wave_col, heights = c(1, 1))
```

## First bout
Spectro + oscillogram
```{r}
bout_1 <- cutw(schroeders, from = 0, to = 0.2, output = "Wave")

spectro(bout_1, scale = FALSE, osc = TRUE, fastdisp = TRUE, grid = FALSE, palette = viridis, collevels = seq(-100, 0 , 5), wl = 512 / 2, colwave = wave_col, heights = c(1, 1))
```

Oscillogram
```{r}

oscillo(bout_1, colwave = wave_col)

```


Oscillogram zooming in
```{r}

oscillo(bout_1, from = 0, to = 0.03, colwave = wave_col)

```


# Amplitude cross-correlation

## Pearson correlation

Try it on the first bout. First get correlation for an amplitude template with the first few samples
```{r}
# make a template from the first fifth of duration of the entire shroeder
template <- bout_1@left[seq_len(length(bout_1@left) / 5)]

cors <- vapply(seq_len(length(bout_1@left) / 4), function(x){
    
        segment <- bout_1@left[x :(x + length(template) - 1)]
    
        cor(template, segment)
}, FUN.VALUE = numeric(1))


plot(cors, type = "l", col = "orange", xlab = "samples", ylab = "Pearson correlation")

```

Then get location of correlation peaks
```{r}
tpks <- fpeaks(cbind(1:length(cors), cors))

# get highest peak
tpks <- tpks[tpks[, 2] > 0.8, ]

# order by sample number
tpks <- tpks[order(tpks[,1]), ]

#remove first peak
tpks <- tpks[-1, ]
```

Then get the mean difference (in samples) between peaks (which is inversely related to the frequency)
```{r}
sample_diffs <- diff(tpks[, 1])

mean_diff <- mean(sample_diffs)

print(round(mean_diff))
```

Now get the start of each period in the sound clip
```{r}
# get envelope peaks
amppk <- fpeaks(cbind(1:length(bout_1@left), abs(bout_1@left) / max(abs(bout_1@left))), plot = FALSE)

# filter highest peaks
amppk <- amppk[amppk[,2] > 0.95, ]

# get the highest one
first_max <- amppk[which.max(amppk[,2]),1] / bout_1@samp.rate

# plot
oscillo(bout_1, from = 0, to = 0.03, colwave = wave_col)

abline(v = first_max + (-100:100 *  (mean(sample_diffs) / bout_1@samp.rate)), col = "red", lty = 2, lwd = 2)

```

# Create a function to spot periods

```{r}

spot_periods <- function(wave, plot = TRUE, abs.amp = TRUE, alpha = 0.5, from = 0, to = 0.03) {

    # make a template from the first fifth of duration of the entire shroeder
    template <- wave@left[seq_len(length(wave@left) / 5)]
    
    cors <- vapply(seq_len(length(wave@left) / 4), function(x) {
        segment <- wave@left[x:(x + length(template) - 1)]
        
        cor(template, segment)
    }, FUN.VALUE = numeric(1))
    
    tpks <- fpeaks(cbind(1:length(cors), cors), plot = FALSE)
    
    # get highest peak
    tpks <- tpks[tpks[, 2] > 0.8,]
    
    # order by sample number
    tpks <- tpks[order(tpks[, 1]),]
    
    #remove first peak
    tpks <- tpks[-1,]
    
    sample_diffs <- diff(tpks[, 1])

    # mean distance in samples between amplitude peaks
   mean_diff <- mean(sample_diffs)
    
    norm_abs <- if (abs.amp) abs(wave@left) / max(abs(wave@left)) else 
        wave@left / max(wave@left)
    
    # get envelope peaks
    amppk <-
        fpeaks(cbind(1:length(wave@left), norm_abs), plot = FALSE)
    
    # filter highest peaks
    amppk <- amppk[amppk[, 2] > 0.95,]
    
    # get the highest one
    first_max <- amppk[which.max(amppk[, 2]), 1] / wave@samp.rate
    
    # get starts of periods
    positions <-
            first_max + (-1000:1000 * (mean_diff / wave@samp.rate))
        
    # filter positions within duration of wave
    positions <- positions[positions > 0 & positions < duration(wave)]

    # plot
    if (plot) {
        osc <- oscillo(wave,
                from = from,
                to = to,
                colwave = wave_col)
        
        abline(
            v = positions,
            col = adjustcolor("blue", alpha.f = alpha),
            lty = 2,
            lwd = 3
        )
        
        points(y = wave@left, x = seq_len(length(wave)) / wave@samp.rate, type = "l", col = wave_col)
    }
    return(positions)
}

```


# Test function

Try function  on other shroeders

5th bout
```{r}

bout_5 <- cutw(schroeders, from = 4.8, to = 5, output = "Wave")

pos <- spot_periods(wave = bout_5, alpha = 0.5)

```

9th bout
```{r}

bout_9 <- cutw(schroeders, from = 10.2, to = 10.4, output = "Wave")

pos <- spot_periods(wave = bout_9, to = 0.02)

```

11th bout
```{r}

bout_11 <- cutw(schroeders, from = 15, to = 15.2, output = "Wave")

pos <- spot_periods(wave = bout_11, to = 0.02)

```

20th bout
```{r}

bout_20 <- cutw(schroeders, from = 31.8, to = 32, output = "Wave")

pos <- spot_periods(wave = bout_20, to = 0.015)

```

---

# Dynamic-time warping

```{r, eval = FALSE}

# make a template from the first fifth of duration of the entire shroeder
template <- bout_1@left[seq_len(length(bout_1@left) / 8)]

dists <- pbapply::pbsapply(cl = 4, seq_len(length(bout_1@left) / 4), function(x){
    
        segment <- bout_1@left[x :(x + length(template) - 1)]
    
        dtw_dist <- dtw::dtwDist(mx = rbind(template, segment))
        return(dtw_dist[1, 2])
})

saveRDS(dists, "./data/processed/dtw_distance_bout_1.RDS")
```


```{r}

dists <- readRDS("./data/processed/dtw_distance_bout_1.RDS")

dists <- dists / max(dists)
sims <-  1 - dists

plot(sims, type = "l", col = "orange", xlab = "samples", ylab = "DTW similarities")

```

Then get location of correlation peaks
```{r}
tpks <- fpeaks(cbind(1:length(sims), sims), threshold = 0.3)
```


Making the time series stationary (`diff()`)
```{r}
stat_sims <- diff(sims)
stat_sims <- stat_sims + abs(min(stat_sims))
stat_sims <- stat_sims / max(stat_sims)

tpks <- fpeaks(cbind(1:length(stat_sims), stat_sims), threshold = 0.98)


# get highest peak
tpks <- tpks[tpks[, 2] > 0.98, ]

sels <- data.frame(sound.files = 1, selec = 1:nrow(tpks), start = tpks[, 1], end = tpks[, 1] + 6, peak = tpks[, 2])

# merge peaks close to each other
sels <- ohun::merge_overlaps(sels, pb = FALSE)
sels$dur <- sels$end - sels$start

tpks <- cbind(sels$start + (sels$dur - 6) / 2, sels$peak)

# order by sample number
tpks <- tpks[order(tpks[,1]), ]

#remove first peak
# tpks <- tpks[-1, ]
```

Then get the difference (in samples) between peaks (which is inversely related to the frequency)
```{r}

sample_diffs <- diff(tpks[, 1])

```


Now get the start of each period in the sound clip
```{r}
# # get envelope peaks
# amppk <- fpeaks(cbind(1:length(bout_1@left), abs(bout_1@left) / max(abs(bout_1@left))), plot = FALSE)
# 
# # filter highest peaks
# amppk <- amppk[amppk[,2] > 0.95, ]
# 
# # get the highest one
# first_max <- amppk[which.max(amppk[,2]),1] / bout_1@samp.rate

# plot
oscillo(bout_1, from = 0, to = 0.03, colwave = wave_col)

abline(v = cumsum(sample_diffs) / bout_1@samp.rate, col = "red", lty = 2, lwd = 2)

```


```{r}

# plot
oscillo(bout_1, from = 0, to = 0.045, colwave = wave_col)

abline(v = cumsum(sample_diffs) / bout_1@samp.rate, col = "red", lty = 2, lwd = 2)

```

<!-- light green box -->

<div class="alert alert-success">

&nbsp; 

# Takeaways

-  Amplitude cross-correlation seems to work fine for getting the periodicity although the position might be off a bit
- Dynamic time warping does also a good job but is much slower
- Dynamic time warping potentially more useful for shroeders in which the periods have variable lengths

</div>

&nbsp;

<!-- '---' adds a gray vertical line -->

---

&nbsp; 
 
 <!-- add packages used, system details and versions  -->
 
<font size="4">Session information</font>

```{r session info, echo=F}

sessionInfo()

```
